# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.

```{r}
dictionary <- readr::read_csv("green_dictionary.csv")
green_data <- readr::read_csv("green_data.csv")
library(tidyverse)
library(psychTools)
library(scales)
```

For your assignment, do the following.

### 1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
  
  > I started by transforming the dataframe into Tibble.

```{r}
as_tibble(green_data)
```


  > I then check the summary of the data 
  
```{r}
green_data %>% 
  summary()
```

  > I noticed that I was running into some issues so I decided to do a quick recode.
  
```{r}
green_data %>% 
mutate(student = recode(student, "1" = "Not_Student", "2" = "Student", .missing = "No_Response", .default = "Unknown")) %>% 
  select(id, student)
```

```{r}
reversed <- dictionary %>% 
  filter(Keying == -1) %>% 
  pull(Item)
``` 

```{r}
green_data %>% 
  mutate(across(all_of(reversed),
                ~ recode(.x, "5" = 1,
                         "4" = 2,
                         "3" = 3,
                         "2" = 4,
                         "1" = 5),
                .names = "{.col}r")) %>% 
  summary()
```

  > Not sure why I was doing all of that work but I decided to check the graph for abnormalities. 
  
```{r}
green_data %>% 
  ggplot() +
  aes(x= comp2) +
  geom_histogram()
```

     Is anything unusual?

#### _Most of the students coding are done in question 2. Refer to question 2 for more information_

> Yes, there is something unusual, the student categories have some listed as NA with all missing values or partially have values. The ones with all NA value throughout the result can be assumed to have been removed while the ones who have partial data are assumed to be unwilling to answer the question. This makes coding much harder as there are currently no sufficient ways to deal with this issue (not as far as the content and knowledge being taught in the clas up until this point).


> Some data are in the -99 which is outside of dictionary Keying and other explainable computation which is affecting the data's characteristic as a whole.

### 2. Compute total scores for the four scales. 
     Recode variables as needed.

 > I started with the basic of what I knew here.
 
```{r}
reversed <- dictionary %>% 
  filter(Keying == -1) %>% 
  pull(Item)
``` 

 > I then recoded everything that is needed to be recoded.
  
```{r}
green_data %>% 
  mutate(across(everything(),
                ~na_if(.x, -99))) %>% 
  summary()
```

```{r}
green_data %>% 
mutate(student = recode(student, "1" = "Not_Student", "2" = "Student", .missing = "No_Response", .default = "Unknown")) %>% 
  select(id, student)
```

```{r}
green_data %>% 
  mutate(across(all_of(reversed),
                ~ recode(.x, "5" = 1,
                         "4" = 2,
                         "3" = 3,
                         "2" = 4,
                         "1" = 5),
                .names = "{.col}r"))
```

  > I decided to make a new data frame with the recoded data.
  
```{r}
new_green_data <- green_data %>% 
  mutate(across(everything(),
                ~na_if(.x, -99))) %>%
  mutate(student = recode(student, "1" = "Not_Student", "2" = "Student", .missing = "No_Response", .default = "Unknown")) %>%
  mutate(across(all_of(reversed),
                ~ recode(.x, "5" = 1,
                         "4" = 2,
                         "3" = 3,
                         "2" = 4,
                         "1" = 5),
                .names = "{.col}r"))
```


  > Since I have the necessary information for the calculation, I decided to make a list of the newly reversed columns into a new variable along with a list of the excluding variables for some of the columns for easier and faster coding process.
  
```{r} 
reversed_columns <- c("green2r", "comp6r", "comp7r", "comp8r", "comp9r", "comp10r", "intel7r", "intel8r", "intel9r", "intel10r", "open7r", "open8r", "open9r", "open10r")
comp_ex <- c("comp6", "comp7", "comp8", "comp9", "comp10")
intel_ex <- c("intel7", "intel8", "intel9", "intel10")
open_ex <- c("open7", "open8", "open9", "open10")
```     

  > I started to do my calculation here with the new exclusion lists and the new data frame.
  
```{r}
new_green_data %>% 
  rowwise() %>%
  mutate(green_tot = mean(c_across(green1:green5 & !green2), na.rm = TRUE)) %>%
  mutate(comp_tot = mean(c_across(comp1:comp10 & !comp_ex), na.rm = TRUE)) %>% 
  mutate(intel_tot = mean(c_across(intel1:intel10 & !intel_ex), na.rm = TRUE)) %>% 
  mutate(open_tot = mean(c_across(open1:open10 & !open_ex),  na.rm = TRUE)) %>% 
  na.omit() %>% 
  select(id, green_tot, comp_tot, intel_tot, open_tot, student)
```

  > 

```{r}
items_data <- new_green_data %>% 
  rowwise() %>%
  mutate(green_tot = mean(c_across(green1:green5 & !green2), na.rm = TRUE)) %>%
  mutate(comp_tot = mean(c_across(comp1:comp10 & !comp_ex), na.rm = TRUE)) %>% 
  mutate(intel_tot = mean(c_across(intel1:intel10 & !intel_ex), na.rm = TRUE)) %>% 
  mutate(open_tot = mean(c_across(open1:open10 & !open_ex),  na.rm = TRUE)) %>% 
  na.omit()
```

```{r}
items_data %>% 
  group_by(student) %>% 
  summarize(green_mean = mean(green_tot, na.rm = TRUE),
            comp_mean = mean(comp_tot, na.rm = TRUE),
            intel_mean = mean(intel_tot, na.rm = TRUE),
            open_mean = mean(open_tot, na.rm = TRUE))
```

```{r}
items_data %>% 
  group_by(student) %>% 
  summarize(green_mean = mean(green_tot, na.rm = TRUE),
            comp_mean = mean(comp_tot, na.rm = TRUE),
            intel_mean = mean(intel_tot, na.rm = TRUE),
            open_mean = mean(open_tot, na.rm = TRUE))
```

### 3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the rescaled variables `*_pomp`.
     
```{r}
items_data %>%
  rowwise() %>% 
  mutate(green_pomp = (100*(green_tot - 1))/4, na.rm = TRUE,
         comp_pomp = (100*(comp_tot - 1))/4, na.rm = TRUE,
         intel_pomp = (100*(intel_tot - 1))/4, na.rm = TRUE,
         open_pomp = (100*(open_tot - 1))/4, na.rm = TRUE) %>% 
  select(id, green_pomp, comp_pomp, intel_pomp, open_pomp)
```

### 4. Make plots that illustrate the distributions of the 4 POMP-scored variables.


### 5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
### 6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
### 7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.
