---
title: "Final_Proj"
author: "Khoa Nguyen"
date: "3/9/2021"
output:  
  html_document: default
  gitub_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggdist)
library(svglite)
library(correlation)
library(splitstackshape)
```

```{r}
if ( !dir.exists( here::here("Final_project", "data_raw") ) ) {
  dir.create( here::here("Final_project", "data_raw", "output", ".R"), recursive = TRUE )
}
```



```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/marriage/divorce.csv"
bothsexes <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/marriage/both_sexes.csv"
men <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/marriage/men.csv"
women <-  "https://raw.githubusercontent.com/fivethirtyeight/data/master/marriage/women.csv"
```

```{r}
Major_data <- read.csv(url,
                       na = c("NA", " ", "-999"))
both_gender <-  read.csv(bothsexes,
                       na = c("NA", " ", "-999"))
men_data <-  read.csv(men,
                       na = c("NA", " ", "-999"))
women_data <-  read.csv(women,
                       na = c("NA", " ", "-999"))
```

```{r}
write.csv(Major_data,
          here::here("Final_project", "data_raw", "Major_data.csv"))
write.csv(both_gender,
          here::here("Final_project", "data_raw", "Both_gender.csv"))
write.csv(men_data,
          here::here("Final_project", "data_raw", "Men_data.csv"))
write.csv(women_data,
          here::here("Final_project", "data_raw", "Women_data.csv"))
```

```{r}
read.csv(
  here::here("Final_Project", "data_raw", "Major_data.csv"))
read.csv(
  here::here("Final_Project", "data_raw", "Both_gender.csv"))
read.csv(
  here::here("Final_Project", "data_raw", "Men_data.csv"))
read.csv(
  here::here("Final_Project", "data_raw", "Women_data.csv"))
```

Summarizing data and observing the marriage trend.

```{r}

m <- men_data %>%
  select(year, starts_with("HS_")) 

m_long <- m %>% pivot_longer(
  cols = -year,
  names_sep = "_",
  names_to = c("edu", "age"),
  values_to = "rate"
)


m_plot <- m_long %>% 
  ggplot(aes(x=year,y=rate, color = age)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Marriage Trend Amongst Different Age Group Over Time for Men.")

m_plot
```

```{r}
ggsave(here::here("Final_project", "output", "figure",
                  "Marriage Trend Amongst Different Age Group Over Time for Men.svg"),
       m_plot,
       height = 6,
       width = 6)
```



```{r}
sum_calculating <- function(.data, variables) {
  
  
  descriptives <- .data %>%
    summarize(across(all_of(variables)), list( mean = mean, sd = sd, min = min, max= max), na.rm = TRUE) %>% 
    pivot_longer(cols = everything(), names_to = c("Education", ".value"), names_sep = "_")

  corr <- .data %>% 
    correlation::correlation(select = variables) %>% 
    summary()

  
  
    
  tab <- full_join(
    descriptives,
    corr,
    by = c("year", "date"
  ))
  
  return(tab)
}

men_ed_data <- men_data %>%
  group_by(year)

women_ed_data <- women_data %>% 
  group_by(year)

education_level <- c("HS_2534", "SC_2534", "BAp_2534", "BAo_2534", "GD_2534")

table_men <- sum_calculating(men_ed_data, education_level)
table_women <-  sum_calculating(women_ed_data, education_level)

table_men
table_women

```

```{r}
sum_for_both_gend <- both_gender %>% 
  group_by(year) %>% 
  summarize(across(c(HS_2534:BAo_2534),
                      list(mean = ~ mean(.x, na.rm= TRUE),
                sd = ~ sd(.x, na.rm = TRUE),
                min = ~ min(.x, na.rm = TRUE),
                max = ~ max(.x, na.rm = TRUE))))
sum_for_both_gend
write.csv(sum_for_both_gend,
          here::here("Final_project", "output", "Sum_both_gend.csv"))
```

```{r}
sum_for_both_gend %>% 
  correlation::correlation(select = c("BAo_2534_mean", "BAp_2534_mean", "SC_2534_mean", "HS_2534_mean")) %>% 
  summary()
  
```

```{r}
tibble_both_gender <- as.tibble(both_gender)
print(tibble_both_gender)
write.csv(tibble_both_gender, 
          here::here("Final_project", "data", "tibble_both_gender.csv"))
```

Bachelor Degree or higher:
```{r}
tibble_both_gender_output1 <- tibble_both_gender %>%
  select(year, date, BAp_2534) %>% 
  group_by(year) %>%
  filter(BAp_2534 == max(BAp_2534))
tibble_both_gender_output1
write.csv(tibble_both_gender_output1,
          here::here("Final_project", "output", "Max_change_in_years_for_higher_education.csv"))
```

As we can see here, the maximum marriage population for people with higher level of education seems to increase over time. This of course is just one variable.

Now, let's look at divorce rate among the population throughout the year by education.

```{r}
Divorce_sum <- Major_data %>% 
  summarize(across(c(HS_3544:BAp_3544),
                      list(mean = ~ mean(.x, na.rm= TRUE),
                sd = ~ sd(.x, na.rm = TRUE),
                min = ~ min(.x, na.rm = TRUE),
                max = ~ max(.x, na.rm = TRUE))))
Divorce_sum
write.csv(Divorce_sum,
          here::here("Final_project", "output", "Divorce_sum.csv"))
```

```{r}
Divorce_trend_output <- Major_data %>%
  select(year, date, BAp_3544) %>% 
  group_by(year) %>%
  filter(BAp_3544 == max(BAp_3544))
Divorce_trend_output
write.csv(Divorce_trend_output,
          here::here("Final_project", "output", "Divorce_trend_output.csv"))
```

As indicated within the data for population that are 35 to 44 years of age, and has a college degree or higher, the trend for divorce rate seems to be increasing by the years when looking at their highest divorce rate for the years (slightly decreasing starting to see significant decreasing in the early 2000s).

Now, let's look at some graphs for the marriage rate versus divorce rate.

I first had to extract out the exact data I want to look at from the divorce dataset and compare it with the marriage one.

```{r}
Divorce_BAp_3544_Data1 <- Major_data %>% 
  mutate(Divorce_BAp_3544 = BAp_3544) %>% 
  select(Divorce_BAp_3544)
Divorce_BAp_3544_Data1
```

```{r}
Marriage_BAp_3544_Data1 <- both_gender %>% 
  mutate(Marriage_BAp_3544 = BAp_3544) %>% 
  select(Marriage_BAp_3544)
Marriage_BAp_3544_Data1
```

I then combined both of them together to make the graph.


```{r}
cbind_BAp_3544_Compar_Data <- cbind(Marriage_BAp_3544_Data1, Divorce_BAp_3544_Data1)
cbind_BAp_3544_Compar_Data
```

```{r}
Rainfall_graph_for_marriage_and_divorce <- cbind_BAp_3544_Compar_Data %>% 
  drop_na(c(Marriage_BAp_3544:Divorce_BAp_3544)) %>% 
  ggplot +
  aes(
    x = Marriage_BAp_3544,
    y = Divorce_BAp_3544,
    fill = Divorce_BAp_3544,
    color = Divorce_BAp_3544
  ) +
  geom_jitter(
    alpha = .5,
    height = .2,
    width = .0,
    na.rm = TRUE
  ) +
  geom_boxplot(
    width = .1,
    color = "Black",
    fill = "transparent"
  ) +
  ggdist::stat_sample_slabinterval(
    alpha = .5,
    height = .2,
    color = "black",
    position = position_nudge(y = .2)
  ) +
  theme_classic() +
  labs(title = "Comparative Data",
       y = NULL,
       x = "BAp_3544"
       ) +
  theme(plot.title.position = "plot") +
  guides(color = guide_none(), fill = guide_none())
Rainfall_graph_for_marriage_and_divorce
```

```{r}
ggsave(here::here("Final_project", "output", "figure",
                  "Rainfall_graph_for_marriage_and_divorce.svg"),
       Rainfall_graph_for_marriage_and_divorce,
       height = 6,
       width = 6)
```

Oh god. This is an abomination. Okay, nothing to see here. Let's go back to regular graph.
